---
title: "Chelicerate coding alignments"
author: "[gwct](https://gwct.github.io/)"
date: "`r format(Sys.time(), '%m/%d/%Y %H:%M:%S %Z')`"
output:
  #html_document:
  rmdformats::robobook:
    highlight: kate
    includes:
      in_header: '../html-chunks/rmd_header.html'
    df_print: paged
    code_folding: hide
    number_sections: true
    toc_depth: 3
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(ggplot2)
library(cowplot)
library(ggbeeswarm)
library(reshape2)
library(ggpubr)
#library(plyr)
library(dplyr)
library(kableExtra)
library(here)
source("../lib/design.r")

```


# Sequence retrieval

Sequence retrieval and alignment of spider orthogroups was done as follows:

1. Extract CDS and peptide sequences for each orthogroup with sequences from at least 4 species (see https://gwct.github.io/spider-wgd/orthologs.html) while accounting for inconsistencies between CDS and peptide sequences (e.g. sometimes the translation of the CDS sequence doesn't match because the peptide sequence has extra X's in it, some CDS have premature stop codons, etc. See below for full accounting).
2. Align CDS sequences with Guidance 2, removing any columns with a score below 0.93 (Guidance default).
3. For each alignment, trim gappy regions by using a sliding window of 3 codons and removing windows that have 2 codons with 2 or more gaps in more than 50% of sequences.
4. After this trimming, for each alignment filter the sequences by removing any that are still made up of >20% gaps.
5. Finally, remove any alignments with sequences from fewer than 4 species or shorter than 33 codons after filtering.
5. Translate to peptide sequence.

# Sequence summary table

```{r seq-summary-table, out.width="50%", fig.align = "center", warning=FALSE}

get_orths_stats = read.csv(here("data", "get-orths-stats.csv"), header=T)

get_orths_stats %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```


## Sequences by species

```{r seq-summary-spec, out.width="50%", fig.align = "center", warning=FALSE}

spec_seq_data_orig = read.csv(here("data", "get-orths-spec-stats.csv"), header=T)

spec_seq_data = select(spec_seq_data_orig, -sequences.total, -abbr)

spec_seq_data_long = melt(spec_seq_data, id.vars=c("species"))

spec_p = ggplot(spec_seq_data_long, aes(x=species, y=value, fill=variable)) +
  geom_bar(position="stack", stat="identity") +
  scale_y_continuous(expand=c(0,0)) +
  xlab("Species") +
  ylab("# of sequences") +
  bartheme() +
  theme(legend.position="bottom") +
  coord_flip()
print(spec_p)


spec_seq_data_orig %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)
```

# Alignment summary table

```{r aln-summary-table, out.width="50%", fig.align = "center", warning=FALSE}

aln_data = read.csv(here("data", "aln-stats-spec4-seq20-site50.log"), sep="\t", header=T, comment.char="#")

aln_summary_data = read.csv(here("data", "aln-stats-summary.csv"), header=T)

aln_summary_data %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

# Alignment length (codons)

Distributions visualized from subset of alignments with length less than 1000 codons. 

- Red points denote average lengths considering the FULL distribution
- Blue points denote median lengths considering the FULL distribution
- Table shows summary statistics from FULL distribution

```{r aln-lens, out.width="50%", fig.align = "center", warning=FALSE}

pre_guid_len = select(aln_data, align, pre.guidance.codon.length)
names(pre_guid_len)[2] = "length"
pre_guid_len$label = "Pre Guidance filtering"

pre_filt_len = select(aln_data, align, pre.codon.aln.length)
names(pre_filt_len)[2] = "length"
pre_filt_len$label = "Pre codon window filtering"

post_filt_len = select(aln_data, align, post.codon.aln.length)
names(post_filt_len)[2] = "length"
post_filt_len$label = "Post filtering"

aln_lens = rbind(pre_guid_len, pre_filt_len, post_filt_len)

aln_means = aln_lens %>% group_by(label) %>% summarize(avg.length=mean(length, na.rm=T), median.length=median(length, na.rm=T), max.length=max(length, na.rm=T))

aln_lens$label = factor(aln_lens$label, levels=c("Pre Guidance filtering", "Pre codon window filtering", "Post filtering"), ordered=T)
aln_lens = subset(aln_lens, length < 1000)

aln_means$label = factor(aln_means$label, levels=c("Pre Guidance filtering", "Pre codon window filtering", "Post filtering"), ordered=T)

aln_lens_p = ggplot(aln_lens, aes(x=label, y=length, group=label)) +
  geom_quasirandom(size=2, width=0.25, alpha=0.1, color="#666666") +
  geom_boxplot(aes(fill=label), outlier.shape=NA, alpha=0.75, width=0.5, color="#666666") +
  geom_point(data=aln_means, aes(x=label, y=avg.length), size=2, color="red") +
  geom_point(data=aln_means, aes(x=label, y=median.length), size=2, color="blue") +
  geom_hline(yintercept=33) +
  ylab("Alignment length (codons)") +
  xlab("") +
  scale_fill_manual(labels=c("Pre Guidance filtering", "Pre codon window filtering", "Post filtering"), values=corecol()) +
  bartheme() +
  theme(legend.position="none", axis.text.x = element_text(angle=15, hjust=1))
print(aln_lens_p)

aln_means[c(3,2,1),] %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

# Average sequence length (codons, ungapped) per alignment

Distributions visualized from subset of alignments with length less than 1000 codons. 

- Red points denote average lengths considering the FULL distribution
- Blue points denote median lengths considering the FULL distribution
- Table shows summary statistics from FULL distribution

Pre = before codon window trimming and gappy sequence removal

```{r ungap-aln-lens, out.width="50%", fig.align = "center", warning=FALSE}

pre_filt_len = select(aln_data, align, pre.avg.nongap.length)
names(pre_filt_len)[2] = "length"
pre_filt_len$label = "Pre filtering"

post_filt_len = select(aln_data, align, post.avg.nongap.length)
names(post_filt_len)[2] = "length"
post_filt_len$label = "Post filtering"

aln_lens = rbind(pre_filt_len, post_filt_len)

aln_means = aln_lens %>% group_by(label) %>% summarize(avg.length=mean(length, na.rm=T), median.length=median(length, na.rm=T), max.length=max(length, na.rm=T))

aln_lens$label = factor(aln_lens$label, levels=c("Pre filtering", "Post filtering"), ordered=T)
aln_lens = subset(aln_lens, length < 1000)

aln_means$label = factor(aln_means$label, levels=c("Pre filtering", "Post filtering"), ordered=T)

aln_lens_p = ggplot(aln_lens, aes(x=label, y=length, group=label)) +
  geom_quasirandom(size=2, width=0.25, alpha=0.1, color="#666666") +
  geom_boxplot(aes(fill=label), outlier.shape=NA, alpha=0.75, width=0.5, color="#666666") +
  geom_point(data=aln_means, aes(x=label, y=avg.length), size=2, color="red") +
  geom_point(data=aln_means, aes(x=label, y=median.length), size=2, color="blue") +
  ylab("Avg. ungapped sequence length (codons)") +
  xlab("") +
  scale_fill_manual(labels=c("Pre filtering", "Post filtering"), values=corecol(pal="wilke")) +
  bartheme() +
  theme(legend.position="none", axis.text.x = element_text(angle=15, hjust=1))
print(aln_lens_p)

aln_means[c(2,1),] %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

# Gappy sites per alignment

Sites where >20% of sequences contain gaps (---)

Recall, the codon window filter removes windows where 2 codons have 2 or more gaps in a 3 codon window in 50% or more of sequences.

Distributions visualized from subset of alignments with fewer than 1000 gappy sites. 

- Red points denote average number of gappy sites considering the FULL distribution
- Blue points denote median number of gappy sites considering the FULL distribution
- Table shows summary statistics from FULL distribution

Pre = before codon window trimming and gappy sequence removal

```{r gappy-sites, out.width="50%", fig.align = "center", warning=FALSE}

pre_filt_sites = select(aln_data, align, pre.gappy.sites)
names(pre_filt_sites)[2] = "gappy.sites"
pre_filt_sites$label = "Pre filtering"

post_filt_sites = select(aln_data, align, post.gappy.sites)
names(post_filt_sites)[2] = "gappy.sites"
post_filt_sites$label = "Post filtering"

aln_gappy_sites = rbind(pre_filt_sites, post_filt_sites)

aln_gappy_means = aln_gappy_sites %>% group_by(label) %>% summarize(avg.gappy.sites=mean(gappy.sites, na.rm=T), median.gappy.sites=median(gappy.sites, na.rm=T), max.gappy.sites=max(gappy.sites, na.rm=T))

aln_gappy_sites$label = factor(aln_gappy_sites$label, levels=c("Pre filtering", "Post filtering"), ordered=T)
aln_gappy_sites = subset(aln_gappy_sites, gappy.sites < 1000)

aln_gappy_means$label = factor(aln_gappy_means$label, levels=c("Pre filtering", "Post filtering"), ordered=T)

aln_gappy_p = ggplot(aln_gappy_sites, aes(x=label, y=gappy.sites, group=label)) +
  geom_quasirandom(size=2, width=0.25, alpha=0.1, color="#666666") +
  geom_boxplot(aes(fill=label), outlier.shape=NA, alpha=0.75, width=0.5, color="#666666") +
  geom_point(data=aln_gappy_means, aes(x=label, y=avg.gappy.sites), size=2, color="red") +
  geom_point(data=aln_gappy_means, aes(x=label, y=median.gappy.sites), size=2, color="blue") +
  ylab("# of gappy sites (codons)") +
  xlab("") +
  scale_fill_manual(labels=c("Pre filtering", "Post filtering"), values=corecol()) +
  bartheme() +
  theme(legend.position="none", axis.text.x = element_text(angle=15, hjust=1))
print(aln_gappy_p)

aln_gappy_means[c(2,1),] %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

# Informative sites (NTs) per alignment

An alignment site is informative if it has at least 2 alleles present in at least 2 sequences.

Distributions visualized from subset of alignments with fewer than 1000 informative sites. 

- Red points denote average number of informative sites considering the FULL distribution
- Blue points denote median number of informative sites considering the FULL distribution
- Table shows summary statistics from FULL distribution

Pre = before codon window trimming and gappy sequence removal

```{r informative-sites-nt, out.width="50%", fig.align = "center", warning=FALSE}

pre_filt_sites = select(aln_data, align, pre.informative.nt.sites)
names(pre_filt_sites)[2] = "informative.sites"
pre_filt_sites$label = "Pre filtering"

post_filt_sites = select(aln_data, align, post.informative.nt.sites)
names(post_filt_sites)[2] = "informative.sites"
post_filt_sites$label = "Post filtering"

aln_informative_sites = rbind(pre_filt_sites, post_filt_sites)

aln_informative_means = aln_informative_sites %>% group_by(label) %>% summarize(avg.informative.sites=mean(informative.sites, na.rm=T), median.informative.sites=median(informative.sites, na.rm=T), max.informative.sites=max(informative.sites, na.rm=T))

aln_informative_sites$label = factor(aln_informative_sites$label, levels=c("Pre filtering", "Post filtering"), ordered=T)
aln_informative_sites = subset(aln_informative_sites, informative.sites < 1000)

aln_informative_means$label = factor(aln_informative_means$label, levels=c("Pre filtering", "Post filtering"), ordered=T)

aln_informative_p = ggplot(aln_informative_sites, aes(x=label, y=informative.sites, group=label)) +
  geom_quasirandom(size=2, width=0.25, alpha=0.1, color="#666666") +
  geom_boxplot(aes(fill=label), outlier.shape=NA, alpha=0.75, width=0.5, color="#666666") +
  geom_point(data=aln_informative_means, aes(x=label, y=avg.informative.sites), size=2, color="red") +
  geom_point(data=aln_informative_means, aes(x=label, y=median.informative.sites), size=2, color="blue") +
  ylab("# of informative sites (nucleotides)") +
  xlab("") +
  scale_fill_manual(labels=c("Pre filtering", "Post filtering"), values=corecol()) +
  bartheme() +
  theme(legend.position="none", axis.text.x = element_text(angle=15, hjust=1))
print(aln_informative_p)

aln_informative_means[c(2,1),] %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)
```

Binning loci by # of informative sites. The largest bin both pre- and post-filter is informative sites = 0.

```{r informative-sites-nt-tables, out.width="50%", fig.align = "center", warning=FALSE}
bins = seq(-1, max(aln_data$pre.informative.nt.sites, na.rm=T), by=1)
bin_labels = seq(0, max(aln_data$pre.informative.nt.sites, na.rm=T), by=1)
pre_counts = hist(aln_data$pre.informative.nt.sites, breaks=bins, include.lowest=T, plot=F)$counts
pre_binned = data.frame("informative.sites"=bin_labels, "loci"=pre_counts)

head(pre_binned) %>% kable(caption="Pre-filter informative sites (NT)") %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)


bins = seq(-1, max(aln_data$post.informative.nt.sites, na.rm=T), by=1)
bin_labels = seq(0, max(aln_data$post.informative.nt.sites, na.rm=T), by=1)
post_counts = hist(aln_data$post.informative.nt.sites, breaks=bins, include.lowest=T, plot=F)$counts
post_binned = data.frame("informative.sites"=bin_labels, "loci"=post_counts)

head(post_binned) %>% kable(caption="Post-filter informative sites (NT)") %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

# Informative sites (codons) per alignment

An alignment site is informative if it has at least 2 alleles present in at least 2 sequences.

Distributions visualized from subset of alignments with fewer than 1000 informative sites. 

- Red points denote average number of informative sites considering the FULL distribution
- Blue points denote median number of informative sites considering the FULL distribution
- Table shows summary statistics from FULL distribution

Pre = before codon window trimming and gappy sequence removal

```{r informative-sites-codon, out.width="50%", fig.align = "center", warning=FALSE}

pre_filt_sites = select(aln_data, align, pre.informative.codon.sites)
names(pre_filt_sites)[2] = "informative.sites"
pre_filt_sites$label = "Pre filtering"

post_filt_sites = select(aln_data, align, post.informative.codon.sites)
names(post_filt_sites)[2] = "informative.sites"
post_filt_sites$label = "Post filtering"

aln_informative_sites = rbind(pre_filt_sites, post_filt_sites)

aln_informative_means = aln_informative_sites %>% group_by(label) %>% summarize(avg.informative.sites=mean(informative.sites, na.rm=T), median.informative.sites=median(informative.sites, na.rm=T), max.informative.sites=max(informative.sites, na.rm=T))

aln_informative_sites$label = factor(aln_informative_sites$label, levels=c("Pre filtering", "Post filtering"), ordered=T)
aln_informative_sites = subset(aln_informative_sites, informative.sites < 1000)

aln_informative_means$label = factor(aln_informative_means$label, levels=c("Pre filtering", "Post filtering"), ordered=T)

aln_informative_p = ggplot(aln_informative_sites, aes(x=label, y=informative.sites, group=label)) +
  geom_quasirandom(size=2, width=0.25, alpha=0.1, color="#666666") +
  geom_boxplot(aes(fill=label), outlier.shape=NA, alpha=0.75, width=0.5, color="#666666") +
  geom_point(data=aln_informative_means, aes(x=label, y=avg.informative.sites), size=2, color="red") +
  geom_point(data=aln_informative_means, aes(x=label, y=median.informative.sites), size=2, color="blue") +
  ylab("# of informative sites (codons)") +
  xlab("") +
  scale_fill_manual(labels=c("Pre filtering", "Post filtering"), values=corecol()) +
  bartheme() +
  theme(legend.position="none", axis.text.x = element_text(angle=15, hjust=1))
print(aln_informative_p)

aln_informative_means[c(2,1),] %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

Binning loci by # of informative sites. The largest bin both pre- and post-filter is informative sites = 0.

```{r informative-sites-codon-tables, out.width="50%", fig.align = "center", warning=FALSE}
bins = seq(-1, max(aln_data$pre.informative.codon.sites, na.rm=T), by=1)
bin_labels = seq(0, max(aln_data$pre.informative.codon.sites, na.rm=T), by=1)
pre_counts = hist(aln_data$pre.informative.codon.sites, breaks=bins, include.lowest=T, plot=F)$counts
pre_binned = data.frame("informative.sites"=bin_labels, "loci"=pre_counts)

head(pre_binned) %>% kable(caption="Pre-filter informative sites (codons)") %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)


bins = seq(-1, max(aln_data$post.informative.codon.sites, na.rm=T), by=1)
bin_labels = seq(0, max(aln_data$post.informative.codon.sites, na.rm=T), by=1)
post_counts = hist(aln_data$post.informative.codon.sites, breaks=bins, include.lowest=T, plot=F)$counts
post_binned = data.frame("informative.sites"=bin_labels, "loci"=post_counts)

head(post_binned) %>% kable(caption="Post-filter informative sites (codons)") %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F)

```

```