---
title: "Chelicerate GRAMPA"
author: "[gwct](https://gwct.github.io/)"
date: "`r format(Sys.time(), '%m/%d/%Y %H:%M:%S %Z')`"
output:
  #html_document:
  rmdformats::robobook:
    highlight: kate
    includes:
      in_header: '../html-chunks/rmd_header.html'
    df_print: paged
    code_folding: hide
    number_sections: true
    toc_depth: 3
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(cowplot)
library(ggtree)
library(viridis)
library(here)
source("../lib/design.r")
source("../lib/get_tree_info.r")

########################################

```

```{r functions}

readGrampa <- function(spec, type, bs, search, label) {
  score_file = here("data", paste0(spec, "spec"), "grampa", paste0(type, "-bs", bs, "-", search, "-scores.txt"))
  count_file = here("data", paste0(spec, "spec"), "grampa", paste0(type, "-bs", bs, "-", search, "-dup-counts.txt"))
  
  score_data = read_tsv(score_file) %>% mutate(rank = row_number()) %>% mutate(label = label) %>% mutate(labeled.tree = paste0(labeled.tree, ";"))
  count_data = read_tsv(count_file)
  # Read the GRAMPA results
  
  names(count_data)[2] = "label"
  
  for(i in levels(as.factor(count_data$mul.tree))) {
    if(i != "0"){
      h_nodes = c()
      for(j in 1:nrow(count_data)){
        if(count_data[j,]$mul.tree == i){
        
          if(grepl("*", count_data[j,]$label, fixed=T)) {
            h_nodes = c(h_nodes, gsub("\\*", "", count_data[j,]$label))
          }
        }
      }
      count_data = count_data %>% mutate(label = ifelse(mul.tree == as.numeric(i) & label %in% h_nodes, paste0(label, "+"), label))
    }
  }
  # This block adds the + to the hybrid clade tips in the dup counts table because GRAMPA doesn't do that...
  
  return(list(score_data, count_data))
}

####################

plotGrampaTrees <- function(data_lowest, dup_counts, title) {
  
  h = corecol(numcol=1, pal="wilke", offset=3)
  l = corecol(numcol=1, offset=3)
  # Colors
  
  tree_fig_list = list()
  tree_info_list = list()
  # A list to store tree figures
  
  first = TRUE
  
  for(i in 1:nrow(data_lowest)){

    cur_tree_num = data_lowest[i,]$mul.tree
    # Get the number of the current tree in the dataset
    
    title_str = paste0("MT-", cur_tree_num)
    if(cur_tree_num == 0) {
     title_str = "ST"
    }
    # Get the label for the plot title
    
    tree = read.tree(text=data_lowest[i,]$labeled.tree)
    tree_to_df_list = treeToDF(tree)
    tree_info = tree_to_df_list[["info"]]
    # Read the current tree
    
    cur_dup_counts = dup_counts %>% filter(mul.tree == cur_tree_num)
    # Get the duplication counts per branch for this tree
    
    tree_info = inner_join(tree_info, cur_dup_counts, by="label") %>% arrange(node)
    # Merge the tree info and dup counts
    
    tree_info$label2 = ifelse(tree_info$node.type == "internal" , tree_info$label, "")
    
    tree_fig = ggtree(tree, size=0.5, ladderize=T, aes(color=dups)) %<+% tree_info +
      scale_color_viridis(name='Duplications', option="C", limits=c(0, max(dup_counts$dups)+100)) +
      #scale_color_continuous(name='Duplications', low=l, high=h, limits=c(0, max(dup_counts$dups)+100)) +
      xlim(0, 15) +
      geom_tiplab(size=2.5) +
      geom_label(aes(x=branch, label=label2), vjust=-.1, size=1.5, color="#920000", fill="transparent", label.size=NA) +
      #geom_nodepoint(color="#666666", alpha=0.85, size=4)
      ggtitle(paste(title_str, " (", data_lowest[i,]$score, ")", sep="")) +
      theme(legend.position="none")
    # Plot the current tree
    #print(tree_fig)
    
    if(first){
      tree_fig = tree_fig + theme(legend.position = "right")
      leg = get_legend(tree_fig)
      tree_fig = tree_fig + theme(legend.position = "none")
      first = FALSE
    }
    
    tree_fig_list[[i]] = tree_fig
    # Add the tree to the tree list
  }
  
  title = plot_grid(NULL, labels=c(title))
  tree_figs_b = plot_grid(plotlist=tree_fig_list, ncol=3)
  tree_figs_leg = plot_grid(tree_figs_b, leg, ncol=2, rel_widths=c(1,0.2))
  tree_figs = plot_grid(title, tree_figs_leg, nrow=2, rel_heights=c(0.1,1))
  
  return(tree_figs)
}

########################################

```

[< Back to project](index.html)

# GRAMPA search limited to horseshoe crab and spider/scorpion clades

## Score distributions

Black dots are the singly-labeled species tree.

Large dots are autopolyploid events.

```{r hcsp-scores, out.width="50%", fig.align = "center", warning=FALSE, fig.height=4}

####################
#rlist = readGrampa(16, "astral-multi", 90, "hcsp", "ASTRAL-Multi")
#multi_hcsp_16 = rlist[[1]]; multi_hcsp_dups_16 = rlist[[2]]

rlist = readGrampa(16, "ballesteros", 90, "hcsp", "Ballesteros")
bal_hcsp_16 = rlist[[1]]; bal_hcsp_dups_16 = rlist[[2]]

rlist = readGrampa(16, "traditional", 90, "hcsp", "Horseshoe crabs sister")
trad_hcsp_16 = rlist[[1]]; trad_hcsp_dups_16 = rlist[[2]]

grampa_data_16 = rbind(bal_hcsp_16, trad_hcsp_16)
grampa_data_st_16 = grampa_data_16 %>% filter(mul.tree == 0)
grampa_data_auto_16 = grampa_data_16 %>% filter(h1.node == h2.node & mul.tree != 0)  

cols = corecol(numcol=3, pal="wilke")
names(cols) = c("ASTRAL-Multi", "Ballesteros", "Horseshoe crabs sister")

####################

scores_fig_16 = ggplot(grampa_data_16, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_16, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_16, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_16$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_16)
# Plot scores

####################

####################
rlist = readGrampa(18, "astral-multi", 90, "hcsp", "ASTRAL-Multi")
multi_hcsp_18 = rlist[[1]]; multi_hcsp_dups_18 = rlist[[2]]

rlist = readGrampa(18, "ballesteros", 90, "hcsp", "Ballesteros")
bal_hcsp_18 = rlist[[1]]; bal_hcsp_dups_18 = rlist[[2]]

rlist = readGrampa(18, "traditional", 90, "hcsp", "Horseshoe crabs sister")
trad_hcsp_18 = rlist[[1]]; trad_hcsp_dups_18 = rlist[[2]]

grampa_data_18 = rbind(multi_hcsp_18, bal_hcsp_18, trad_hcsp_18)
grampa_data_st_18 = grampa_data_18 %>% filter(mul.tree == 0)
grampa_data_auto_18 = grampa_data_18 %>% filter(h1.node == h2.node & mul.tree != 0)  

cols = corecol(numcol=3, pal="wilke")
names(cols) = c("ASTRAL-Multi", "Ballesteros", "Horseshoe crabs sister")

####################

scores_fig_18 = ggplot(grampa_data_18, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_18, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_18, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_18$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_18)
# Plot scores

####################

####################

rlist = readGrampa(19, "astral-multi", 90, "hcsp", "ASTRAL-Multi")
multi_hcsp_19 = rlist[[1]]; multi_hcsp_dups_19 = rlist[[2]]

rlist = readGrampa(19, "ballesteros", 90, "hcsp", "Ballesteros")
bal_hcsp_19 = rlist[[1]]; bal_hcsp_dups_19 = rlist[[2]]

rlist = readGrampa(19, "traditional", 90, "hcsp", "Horseshoe crabs sister")
trad_hcsp_19 = rlist[[1]]; trad_hcsp_dups_19 = rlist[[2]]

grampa_data_19 = rbind(multi_hcsp_19, bal_hcsp_19, trad_hcsp_19)
grampa_data_st_19 = grampa_data_19 %>% filter(mul.tree == 0)
grampa_data_auto_19 = grampa_data_19 %>% filter(h1.node == h2.node & mul.tree != 0)  

####################

scores_fig_19 = ggplot(grampa_data_19, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_19, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_19, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_19$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_19)
# Plot scores

####################

leg = get_legend(scores_fig_19)

scores_fig_top = plot_grid(scores_fig_16 + theme(legend.position="none"), 
                           scores_fig_18 + theme(legend.position="none"), 
                           scores_fig_19 + theme(legend.position="none"), 
                           ncol=3, labels=c("16 species", "18 species", "19 species"))
scores_fig = plot_grid(scores_fig_top, leg, nrow=2, rel_heights=c(1,0.1))
print(scores_fig)

```

## Lowest scoring trees - ASTRAL-Multi

```{r cds-apro-hcsp-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

#amulti_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_hcsp_dups_16, "16 species")
amulti_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_hcsp_dups_18, "18 species")
amulti_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_hcsp_dups_19, "19 species")

print(amulti_trees_18)
print(amulti_trees_19)
# Combine and render figs

```


```{r, eval=F} 
<details><summary>Click here to see a breakdown the above command</summary>
<p>

multi_hcsp_dups_19 %>% kable() %>% kable_styling(bootstrap_options=c("striped", "condended", "responsive"), full_width=F, position="center", fixed_thead=T) %>% scroll_box(width="400px", height="500px")

</p>
</details>
</br>
```

## Lowest scoring trees - Ballesteros

```{r ballesteros-hcsp-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

bal_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_hcsp_dups_16, "16 species")
bal_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_hcsp_dups_18, "18 species")
bal_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_hcsp_dups_19, "19 species")

print(bal_trees_16)
print(bal_trees_18)
print(bal_trees_19)
# Combine and render figs

```


## Lowest scoring trees - Horseshoe crabs sister

```{r traditional-hcsp-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

trad_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_hcsp_dups_16, "16 species")
trad_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_hcsp_dups_18, "18 species")
trad_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_hcsp_dups_19, "19 species")

print(trad_trees_16)
print(trad_trees_18)
print(trad_trees_19)
# Combine and render figs

```

# Full GRAMPA search

## Score distributions

Black dots are the singly-labeled species tree.

Large dots are autopolyploid events.

```{r full-scores, out.width="50%", fig.align = "center", warning=FALSE, fig.height=4}

####################

rlist = readGrampa(16, "astral-multi", 90, "full", "ASTRAL-Multi")
multi_full_16 = rlist[[1]]; multi_full_dups_16 = rlist[[2]]

rlist = readGrampa(16, "ballesteros", 90, "full", "Ballesteros")
bal_full_16 = rlist[[1]]; bal_full_dups_16 = rlist[[2]]

rlist = readGrampa(16, "traditional", 90, "full", "Horseshoe crabs sister")
trad_full_16 = rlist[[1]]; trad_full_dups_16 = rlist[[2]]

grampa_data_16 = rbind(multi_full_16, bal_full_16, trad_full_16)
grampa_data_st_16 = grampa_data_16 %>% filter(mul.tree == 0)
grampa_data_auto_16 = grampa_data_16 %>% filter(h1.node == h2.node & mul.tree != 0)  

####################

scores_fig_16 = ggplot(grampa_data_16, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_16, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_16, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_16$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_16)
# Plot scores

####################

####################

rlist = readGrampa(18, "astral-multi", 90, "full", "ASTRAL-Multi")
multi_full_18 = rlist[[1]]; multi_full_dups_18 = rlist[[2]]

rlist = readGrampa(18, "ballesteros", 90, "full", "Ballesteros")
bal_full_18 = rlist[[1]]; bal_full_dups_18 = rlist[[2]]

rlist = readGrampa(18, "traditional", 90, "full", "Horseshoe crabs sister")
trad_full_18 = rlist[[1]]; trad_full_dups_18 = rlist[[2]]

grampa_data_18 = rbind(multi_full_18, bal_full_18, trad_full_18)
grampa_data_st_18 = grampa_data_18 %>% filter(mul.tree == 0)
grampa_data_auto_18 = grampa_data_18 %>% filter(h1.node == h2.node & mul.tree != 0)  

####################

scores_fig_18 = ggplot(grampa_data_18, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_18, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_18, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_18$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_16)
# Plot scores

####################

####################

rlist = readGrampa(19, "astral-multi", 90, "full", "ASTRAL-Multi")
multi_full_19 = rlist[[1]]; multi_full_dups_19 = rlist[[2]]

rlist = readGrampa(19, "ballesteros", 90, "full", "Ballesteros")
bal_full_19 = rlist[[1]]; bal_full_dups_19 = rlist[[2]]

rlist = readGrampa(19, "traditional", 90, "full", "Horseshoe crabs sister")
trad_full_19 = rlist[[1]]; trad_full_dups_19 = rlist[[2]]

grampa_data_19 = rbind(multi_full_19, bal_full_19, trad_full_19)
grampa_data_st_19 = grampa_data_19 %>% filter(mul.tree == 0)
grampa_data_auto_19 = grampa_data_19 %>% filter(h1.node == h2.node & mul.tree != 0)  

####################

scores_fig_19 = ggplot(grampa_data_19, aes(x=rank, y=score, color=label)) +
  geom_point(size=2, alpha=0.33, show.legend=F) +
  geom_point(data=grampa_data_st_19, aes(x=rank, y=score), size=3, alpha=0.75, color="#333333", fill="#999999") +
  geom_point(data=grampa_data_auto_19, aes(x=rank, y=score, color=label), size=5, alpha=0.5) +
  scale_x_continuous(limits=c(-10,max(grampa_data_19$rank)+10)) +
  scale_color_manual(values=cols) +
  xlab("Rank") +
  ylab("GRAMPA score") +
  bartheme() +
  theme(legend.position="bottom")
#print(scores_fig_19)
# Plot scores

####################

leg = get_legend(scores_fig_19)

scores_fig_top = plot_grid(scores_fig_16 + theme(legend.position="none"), 
                           scores_fig_18 + theme(legend.position="none"), 
                           scores_fig_19 + theme(legend.position="none"), 
                           ncol=3, labels=c("16 species", "18 species", "19 species"))
scores_fig = plot_grid(scores_fig_top, leg, nrow=2, rel_heights=c(1,0.1))
print(scores_fig)

```

## Lowest scoring trees - ASTRAL-Multi

```{r cds-apro-full-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

amulti_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_full_dups_16, "16 species")
amulti_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_full_dups_18, "18 species")
amulti_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "ASTRAL-Multi") %>% slice_head(n=6), multi_full_dups_19, "19 species")

print(amulti_trees_16)
print(amulti_trees_18)
print(amulti_trees_19)
# Combine and render figs

```

## Lowest scoring trees - Ballesteros

```{r ballesteros-full-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

bal_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_full_dups_16, "16 species")
bal_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_full_dups_18, "18 species")
bal_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "Ballesteros") %>% slice_head(n=6), bal_full_dups_19, "19 species")

print(bal_trees_16)
print(bal_trees_18)
print(bal_trees_19)
# Combine and render figs

```

## Lowest scoring trees - Horseshoe crabs sister

```{r traditional-full-trees, out.width="65%", fig.align = "center", warning=FALSE, fig.height=6}

trad_trees_16 = plotGrampaTrees(grampa_data_16 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_full_dups_16, "16 species")
trad_trees_18 = plotGrampaTrees(grampa_data_18 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_full_dups_18, "18 species")
trad_trees_19 = plotGrampaTrees(grampa_data_19 %>% filter(label == "Horseshoe crabs sister") %>% slice_head(n=6), trad_full_dups_19, "19 species")

print(trad_trees_16)
print(trad_trees_18)
print(trad_trees_19)
# Combine and render figs

```


