---
title: "Chelicerate ortholog clustering stats"
author: "[gwct](https://gwct.github.io/)"
date: "`r format(Sys.time(), '%m/%d/%Y %H:%M:%S %Z')`"
output:
  #html_document:
  rmdformats::robobook:
    highlight: kate
    includes:
      in_header: '../html-chunks/rmd_header.html'
    df_print: paged
    code_folding: hide
    number_sections: true
    toc_depth: 3
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(ggplot2)
library(dplyr)
library(ggrepel)
library(here)
library(reshape2)
library(kableExtra)
source("../lib/design.r")
```

[< Back to samples](index.html)

Sequences and annotations from each genome were downloaded from various databases in 10.21

```{r read-data}

#infilename = here("..", "..", "data", "chelicerate_genomes-gwct.csv")
infilename = here("data", "fastortho-summary.csv")
in_data = read.csv(infilename, header=T, comment.char="#")

```

## Number of clusters

```{r num-clusters, out.width="50%", fig.align="center"}

num_clusters = in_data %>% group_by(inflation) %>% summarize("num.clusters"=n())

num_clusters_p = ggplot(num_clusters, aes(x=inflation, y=num.clusters)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(num_clusters$num.clusters)+1000)) +
  ylab("# clusters") +
  xlab("Inflation value") +
  bartheme()
print(num_clusters_p)

```

## Max cluster size

```{r max-size, out.width="50%", fig.align="center"}

max_size = in_data %>% group_by(inflation) %>% summarize("max.size"=max(genes))

max_size_p = ggplot(max_size, aes(x=inflation, y=max.size)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(max_size$max.size)+100)) +
  ylab("Max cluster size") +
  xlab("Inflation value") +
  bartheme()
print(max_size_p)

```

## Mean cluster size

```{r mean-size, out.width="50%", fig.align="center"}

mean_size = in_data %>% group_by(inflation) %>% summarize("mean.size"=mean(genes))

mean_size_p = ggplot(mean_size, aes(x=inflation, y=mean.size)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(mean_size$mean.size)+10)) +
  ylab("Mean cluster size") +
  xlab("Inflation value") +
  bartheme()
print(mean_size_p)

```

## Num. single gene clusters

```{r single-gene, out.width="50%", fig.align="center"}

single_genes = subset(in_data, genes==1)
single_genes = single_genes %>% group_by(inflation) %>% summarize("single.genes"=n())


single_gene_p = ggplot(single_genes, aes(x=inflation, y=single.genes)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(single_genes$single.genes)+100)) +
  ylab("# single gene clusters") +
  xlab("Inflation value") +
  bartheme()
print(single_gene_p)

```

## % single gene clusters

```{r perc-single-gene, out.width="50%", fig.align="center"}

#single_genes = subset(in_data, genes==1)
#single_genes = single_genes %>% group_by(inflation) %>% summarize("single.genes"=n())

perc_single_genes = merge(num_clusters, single_genes, by="inflation")
perc_single_genes$perc.single.genes = perc_single_genes$single.genes / perc_single_genes$num.clusters * 100

perc_single_gene_p = ggplot(perc_single_genes, aes(x=inflation, y=perc.single.genes)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(perc_single_genes$perc.single.genes)+100)) +
  ylab("% single gene clusters") +
  xlab("Inflation value") +
  bartheme()
print(perc_single_gene_p)

```

## Single copy clusters

The number of clusters with 1 gene present in each of the 16 species

```{r single-copy, out.width="50%", fig.align="center"}

single_copy = subset(in_data, genes==16 & taxa==16)
single_copy = single_copy %>% group_by(inflation) %>% summarize("single.copy"=n())

single_copy_p = ggplot(single_copy, aes(x=inflation, y=single.copy)) +
  geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(single_copy$single.copy)+100)) +
  ylab("# single copy clusters") +
  xlab("Inflation value") +
  bartheme()
print(single_copy_p)

```


## Cluster size distributions by inflation parameter

```{r cluster-dists, out.width="66%", fig.align="center"}

dist_data = subset(in_data, genes <= 30)

dist_p = ggplot(dist_data, aes(x=genes)) +
  geom_histogram(bins=30, fill=corecol(pal="wilke", numcol=1, offset=1), color="#ececec") +
  facet_wrap(~inflation) +
  #geom_line(color=corecol(pal="wilke", numcol=1, offset=2)) +
  #geom_point(size=4, color=corecol(pal="wilke", numcol=1, offset=2)) +
  #scale_y_continuous(expand=c(0,0), limits=c(0,max(single_copy$single.copy)+100)) +
  ylab("# of clusters") +
  xlab("Cluster size") +
  bartheme()
print(dist_p)

```

## All stats table

```{r table, out.width="66%", fig.align="center"}

t_data = list(num_clusters, max_size, mean_size, single_genes, perc_single_genes, single_copy)
t_data = Reduce(function(x, y) merge(x, y, all=TRUE), t_data)

t_data %>% kable(caption="Cluster statistics") %>% kable_styling(bootstrap_options=c("striped", "condensed", "responsive"), full_width=F)

```

[< Back to samples](index.html)